---
title: "Project2"
output: html_document
date: "2022-08-08"
---

Importing the data and necessary modules
```{r setup, include=FALSE}
library(tidyverse)
library(GGally)
library(reshape2)
library(gridExtra)
house <- read.csv('houses_clean.csv', header=TRUE)
```

Removing X Column and Splitting Training and Test Data
```{r}
house <- subset(house, select = -c(X))
set.seed(6021)
sample<-sample.int(nrow(house), floor(.80*nrow(house)), replace = F)
train<-house[sample, ]
test<-house[-sample, ]
```

Summary Statistics for Training Data
```{r}
summary(train)
options(scipen=10000)
```

Standard Deviation for Training Data
```{r}
apply(train,2,sd)
```


Histogram for Sale Price
```{r}
train %>% 
  ggplot(aes(x=price))+
  geom_histogram()+
  ggtitle('Sale Price')+
  xlab('Price (in USD)')+
  ylab('Count')
```

Zoomed in Histogram for Sale Price
```{r}
train %>% 
  filter(price <= 4000000) %>% 
  ggplot(aes(x=price))+
  geom_histogram()+
  ggtitle('Sale Price')+
  xlab('Price (in USD)')+
  ylab('Count')
```

Zoomed in Histogram of Price for houses in Seattle and Outside of Seattle
```{r}
train %>% 
  filter(price <= 3000000) %>% 
  ggplot(aes(x=price))+
  geom_histogram(aes(color=seattle, alpha=0.5))
```
Scatterplot of Bedrooms vs Sale Price
```{r}
bedsale <- train %>% 
  ggplot(aes(x=bedrooms, y=price))+
  geom_point(alpha=0.1)+
  ggtitle('Bedrooms and Sale Price')+
  xlab('Number of Bathrooms')+
  ylab('Price (in USD)')
```

Scatterplot of Bathrooms vs Sale Price
```{r}
bathsale <- train %>%
  ggplot(aes(x=bathrooms, y=price))+
  geom_point(alpha=0.1)+
  ggtitle('Bathrooms and Sale Price')+
  xlab('Number of Bedrooms')+
  ylab('Price (in USD)')+
  geom_smooth(method='lm', se=FALSE)
```

Scatterplot of Living Space (in square feet) vs Sale Price
```{r}
livingsale <- train %>% 
  ggplot(aes(x=sqft_living, y=price))+
  geom_point(alpha=0.1)+
  ggtitle('Living Space and Sale Price')+
  xlab('Living Space (in sq. ft.)')+
  ylab('Price (in USD)')+
  geom_smooth(method='lm', se=FALSE)
```

Scatterplot of Lot Size (in square feet) vs Sale Price
```{r}
lotsale <- train %>%
  ggplot(aes(x=sqft_lot, y=price))+
  geom_point(alpha=0.1)+
  ggtitle('Lot Size and Sale Price')+
  xlab('Lot Size (in sq. ft.)')+
  ylab('Price (in USD)')+
  geom_smooth(method='lm', se=FALSE)
```

Grid of Previous 4 Graphs
```{r}
grid.arrange(bedsale,bathsale,livingsale, lotsale, ncol=2)
```

Boxplot of Number of Floors vs Sale Price
```{r}
floorsale <- train %>% 
  ggplot(aes(x=as.factor(floors), y=price))+
  geom_boxplot()+
  ggtitle('Floors and Sale Price')+
  xlab('Number of Floors')+
  ylab('Price (in USD)')

```

Boxplot of Waterfront View vs Sale Price
```{r}
watersale <- train %>% 
  ggplot(aes(x=factor(waterfront, labels=c('No', 'Yes')), y=price))+
  geom_boxplot()+
  ggtitle('Waterfronts and Sale Price')+
  xlab('Is there a Waterfront View?')+
  ylab('Price (in USD)')

```

Boxplot of View Quality vs Sale Price
```{r}
viewsale <- train %>% 
  ggplot(aes(x=as.factor(view), y=price))+
  geom_boxplot()+
  ggtitle('View Quality and Sale Price')+
  xlab('View Rating')+
  ylab('Price (in USD)')

```

Scatterplot of Year Built vs Sale Price
```{r}
yrsale <- train %>%
  ggplot(aes(x=yr_built, y=price))+
  geom_point(alpha=0.1)+
  ggtitle('Year Built and Sale Price')+
  xlab('Year Built')+
  ylab('Price (in USD)')+
  geom_smooth(method='lm', se=FALSE)

```

Grid of previous 4 Graphs
```{r}
grid.arrange(floorsale,watersale,viewsale, yrsale, ncol=2)
```
Scatterplot of Neighbor's Living Space vs Sale Price
```{r}
living15sale <- train %>%
  ggplot(aes(x=sqft_living15, y=price))+
  geom_point(alpha=0.1)+
  ggtitle("Neighbors' Living Space and Price")+
  xlab("Neighbors' Living Space (in sq. ft.)")+
  ylab('Price (in USD)')+
  geom_smooth(method='lm', se=FALSE)

```

Scatterplot of Neighbor's Lot Size vs Sale Price
```{r}
lot15sale <- train %>%
  ggplot(aes(x=sqft_lot15, y=price))+
  geom_point(alpha=0.1)+
  ggtitle("Neighbors' Lot Size and Sale Price")+
  xlab("Neighbors' Lot Size (in sq. ft.)")+
  ylab('Price (in USD)')+
  geom_smooth(method='lm', se=FALSE)

```

Boxplot of Seattle vs Sale Price
```{r}
seattlesale <- train %>% 
  ggplot(aes(x=factor(seattle, labels=c('No', 'Yes')), y=price))+
  geom_boxplot()+
  ggtitle('Seattle and Sale Price')+
  xlab('Is the House in Seattle')+
  ylab('Price (in USD)')

```


Boxplot of Renovated vs Sale Price
```{r}
renovatedsale <- train %>% 
  ggplot(aes(x=factor(renovated, labels=c('No', 'Yes')), y=price))+
  geom_boxplot()+
  ggtitle('Renovated Houses and Sale Price')+
  xlab('Has the House Ever Been Renovated?')+
  ylab('Price (in USD)')

```

Grid of Seattle and Renovated Houses Plots
```{r}
grid.arrange(seattlesale,renovatedsale, ncol=2)
```

Grid of Neighbor's Property vs Price
```{r}
grid.arrange(living15sale,lot15sale, ncol=2)
```

Correlation Matrix
```{r}
#cor(subset(house, select = c(price_per_sqft, price, sqft_living, yr_built, renovated, view)))
cormelt <- melt(cor(train))

head(cormelt)
ggplot(data = cormelt, aes(x=Var1, y=Var2, fill=value))+
  geom_tile()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_text(aes(label = round(value, 2)), color = "white", size = 2) +
  guides(fill = guide_colourbar(title = 'Correlation'))+
  coord_fixed()
```
